---
title: "20251101_Baseline_Outcome"
author: "Yu-Chun Huang"
output: html_document
date: "`r format(Sys.time(), '%Y-%m-%d')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Background

-   **Author**: `<Yu-Chun Huang>`
-   **Created At**: `<2025-11-01>`
-   **Research Motivation and Context (why are we interested in the findings?)：**
-   **Main Findings and Takeaways：**
-   **Future Direciton：**

```{r}
library(readxl)
library(tidyverse)
library(janitor)
library(lubridate)
library(lmtest)
library(ranger)
library(broom)
library(gt)
library(glue)
library(ggplot2)
```

```{r}
# import data
austin <- read.csv("/Users/nnychhi/Desktop/碩二/austin.csv")

# data clean up
austin_clear <- austin %>%
  # drop irrelevant variable
  select(-c("Name", "Found.Location", "Sex.upon.Intake", "Age.upon.Intake", 
            "Date.of.Birth", "Outcome.Subtype", "Age.upon.Outcome")) %>%
  # rename in lower case
  rename_with(tolower) %>%
  # filter for animal and outcome type 
  filter(animal.type %in% c("Cat", "Dog"), outcome.type == "Adoption") %>%
  mutate(animal.type = factor(animal.type, levels = c("Dog", "Cat"))) %>% # 0: dog
  # standardized name form
  rename(intake.datetime = intake_datetime,
         outcome.datetime = outcome_datetime) %>%
  # calculate duration of stay
  mutate(start.date = ymd_hms(intake.datetime),
         end.date = ymd_hms(outcome.datetime),
         stay.day.round = round(as.numeric(difftime(end.date, start.date, units = "days")), 2),
         stay.day = floor(as.numeric(difftime(end.date, start.date, units = "days")))) %>%
  # adjustment
  select(-c("outcome.type", "start.date", "end.date"))%>%
  # variable redefined: sex_upon_outcome
  mutate(
    sterilized = case_when(
      grepl("Neutered|Spayed", sex.upon.outcome, ignore.case = TRUE) ~ "Sterilized",
      grepl("Intact", sex.upon.outcome, ignore.case = TRUE) ~ "Intact",
      TRUE ~ "Unknown",
    ),
    sex = case_when(
      sex.upon.outcome == "Unknown" ~ "Unknown",
      TRUE ~ str_to_title(word(sex.upon.outcome, -1))  # pick last word：Male / Female
    ),
    sterilized = factor(sterilized, levels = c("Intact", "Sterilized", "Unknown")), # 0:未絕育
    sex = factor(sex, levels = c("Male", "Female", "Unknown")) # 0: 雄性
  ) %>%
  select(-sex.upon.outcome) %>%
  # variable redefined: breed
  mutate(is_mixed = if_else(str_detect(breed, "Mix|/|-|&"), "Mixed", "Purebred"),
         is_mixed = factor(is_mixed, levels = c("Purebred", "Mixed")) # 0: 純種
         ) %>%
  # variable redefined: color
  mutate(
    color_type = case_when(
      str_detect(color, "(?i)Tabby|Brindle|Tortie|Calico|Merle|Point|Tick|Spotted|Speckled") ~ "Patterned",
      str_count(color, "/|&| ") == 0 ~ "Solid",
      str_count(color, "/|&| ") == 1 ~ "Bicolor",
      str_count(color, "/|&| ") >= 2 ~ "Tricolor+",
      TRUE ~ "Unknown"
    ),
    color_type = factor(color_type, levels = c("Solid", "Bicolor", "Tricolor+", "Patterned")) # 0: 純色
  ) %>%
  # adjust variable name form
  rename_with(~ str_replace_all(.x, "\\.", "_"))
  
# further selection for analysis 
# (austin_clear remains some vars for future analysis)
austin_for_analysis <- austin_clear %>%
  select(-c("intake_datetime", "intake_type", "intake_condition", "stay_day",
            "breed", "color", "outcome_datetime")) %>%
  mutate(log_stay_day_round = log1p(stay_day_round)) %>%
  # drop incomplete observations
  filter(
    if_all(
      c(sex, sterilized, color_type),
      ~ . != "Unknown" & !is.na(.)
    )
  )
# print first 10 obs
head(austin_for_analysis, 10)
```

```{r}
# Before conducting the analysis, report summary statistics for all variables 
# that will be used in this notebook:
#   - For numeric variables: mean, min, 25th percentile, median, 75th percentile, 
#     max, standard deviation, and number of observations.
#   - For categorical variables: number of observations and the counts/percentages 
#     of the ten most frequent categories.
# Conclude this block with the complete summary statistics.
```

### The actual analysis starts below

#### OLS estimation

```{r}
# baseline analysis: OLS + RF
## OLS ---
lm_baseline <- lm(log_stay_day_round ~ animal_type + sterilized + sex + is_mixed 
                  + color_type, data = austin_for_analysis)

lm_baseline_summary <- tidy(lm_baseline, conf.int = TRUE)

vc_hc1_lm_baseline <- sandwich::vcovHC(lm_baseline, type = "HC1")
tidy_hc1 <- broom::tidy(lm_baseline, conf.int = TRUE, vcov = vc_hc1_lm_baseline)

# star for estimated coefficients
star <- function(p){
  ifelse(p < 0.01, "***",
         ifelse(p < 0.05, "**",
                ifelse(p < 0.10, "*", "")))
}

# plot: estimated coefficient
coef_plot <- tidy_hc1 %>%
  filter(term != "(Intercept)") %>%
  mutate(term = recode(term,
                       "animal_typeCat" = "Animal: Cat",
                       "sterilizedSterilized" = "Sterilized",
                       "sexFemale" = "Sex: Female",
                       "is_mixedMixed" = "Breed: Mixed",
                       "color_typeBicolor" = "Color: Bicolor",
                       "color_typeTricolor+" = "Color: Tricolor+",
                       "color_typePatterned" = "Color: Patterned"
  ))

ggplot(coef_plot, aes(x = term, y = estimate)) +
  geom_point(size=3, color="steelblue") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color="steelblue") +
  geom_hline(yintercept = 0, linetype="dashed", color="gray50") +
  coord_flip() +
  labs(
    title = "Effect of Animal Characteristics on Days to Adoption (log(1 + days))",
    subtitle = "OLS coefficients with 95% confidence intervals (HC1 robust SE)",
    x = "",
    y = "Coefficient"
  )
```

```{r}
# result table preparation
tbl_coef <- tidy_hc1 %>%
  filter(term != "(Intercept)") %>%
  mutate(
    term = recode(term,
                  "animal_typeCat"     = "Animal: Cat",
                  "sterilizedSterilized" = "Sterilized",
                  "sexFemale"          = "Sex: Female",
                  "is_mixedMixed"      = "Breed: Mixed",
                  "color_typeBicolor"  = "Color: Bicolor",
                  "color_typeTricolor+"= "Color: Tricolor+",
                  "color_typePatterned"= "Color: Patterned"
    ),
    sig  = star(p.value),
    est_with_star = ifelse(sig=="", 
                           sprintf("%.3f", estimate),
                           sprintf("%.3f<sup>%s</sup>", estimate, sig)),
    ci_col = sprintf("[%.3f, %.3f]", conf.low, conf.high)
  ) %>%
  transmute(
    Term = term,
    Estimate = est_with_star,
    `Std. Error (HC1)` = sprintf("%.3f", std.error),
    `95% CI` = ci_col,
    `p-value` = sprintf("%.3f", p.value)
  )

# add outcome variable mean, r square
g_stats <- glance(lm_baseline)
y_mean  <- mean(austin_for_analysis$log_stay_day_round, na.rm = TRUE)

# result table
result_tbl <- tbl_coef %>%
  gt() %>%
  tab_header(
    title = md("**OLS results with HC1 robust standard errors**"),
    subtitle = md(glue("Outcome mean (log(1 + days)) = **{sprintf('%.3f', y_mean)}**"))
  ) %>%
  fmt_markdown(columns = "Estimate") %>%
  cols_label(
    Term = "Variable",
    Estimate = "Estimate",
    `Std. Error (HC1)` = "Std. Error (HC1)",
    `95% CI` = "95% CI",
    `p-value` = "p-value"
  ) %>%
  cols_align(
    align = "left",
    columns = vars(Term)
  ) %>%
  cols_align(
    align = "right",
    columns = c("Estimate", "Std. Error (HC1)", "95% CI", "p-value")
  ) %>%
  tab_source_note(
    source_note = md(
      glue("**R²** = {sprintf('%.3f', g_stats$r.squared)}; ",
           "**Adj. R²** = {sprintf('%.3f', g_stats$adj.r.squared)}; ",
           "**N** = {g_stats$nobs}")
    )
  ) %>%
  tab_source_note(
    source_note = md(
      "**Base groups:** Animal = Dog; Sterilization = Intact; Sex = Male; Breed = Purebred; Color = Solid."
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(size = "smaller", color = "gray40", align = "left")
    ),
    locations = cells_source_notes()
  ) %>%
  opt_row_striping()

result_tbl
```

#### Random forest estimation

We apply random forest model to estimate the importance of each variables, and the result can be compared with the result of OLS.

To conduct random forest estimation. First, we find the best number of covariates randomly selected at each split. (denoted by m in the lecture note)

```{r}
## RF ---
set.seed(5166)

# find the best m (number of covariates randomly selected at each split)
p <- 5
mtry_values <- 1:p

m_results <- data.frame(mtry = mtry_values, OOB_MSE = NA)

 # test for different m
for (i in seq_along(mtry_values)) {
  rf_temp <- ranger(
    log_stay_day_round ~ animal_type + sterilized + sex + is_mixed + color_type,
    data = austin_for_analysis,
    num.trees = 500,
    mtry = mtry_values[i],
    importance = "impurity"
  )
  m_results$OOB_MSE[i] <- rf_temp$prediction.error
}

 # best m: 3
print(m_results)
#best_mtry <- m_results$mtry[which.min(m_results$OOB_MSE)]
#cat("Best mtry =", best_mtry, "\n")
```
We can see that the OOB_MSE is lowest when m = 3.

So we then conduct random forest and set m = 3.
The result shows below:
```{r}
# RF model
rf_baseline <- ranger(
  log_stay_day_round ~ animal_type + sterilized + sex + is_mixed + color_type,
  data = austin_for_analysis,
  num.trees = 500,
  mtry = 3,
  importance = "permutation" 
  # use permutation since we focus on importance
)

rf_baseline$variable.importance

# rf importance plot
rf_importance <- as.data.frame(rf_baseline$variable.importance) %>%
  rownames_to_column("Variable") %>%
  rename(Importance = 2)

# adjust variable names
as_ols_names <- c(
  animal_type = "Animal type",
  sterilized  = "Sterilized",
  sex         = "Sex",
  is_mixed    = "Breed",
  color_type  = "Color pattern"
)

display_order <- c("Animal type","Sterilized","Sex","Breed","Color pattern")

rf_importance <- rf_importance %>%
  mutate(
    Variable = recode(Variable, !!!as_ols_names),
    Variable = factor(Variable, levels = rev(display_order))
  )

ggplot(rf_importance, aes(x = Variable, y = Importance)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title    = "Variable Importance from Random Forest",
    subtitle = "Permutation-based importance (500 trees, mtry = 3)",
    x = "",
    y = "Importance (Increase in prediction error)",
    caption = paste(
      "Base groups (OLS): Animal = Dog; Sterilization = Intact; Sex = Male;",
      "Breed = Purebred; Color = Solid.\n"
    )
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.caption = element_text(
      hjust = 0, color = "gray35", size = 10,
      margin = ggplot2::margin(t = 8, unit = "pt") 
    ),
    plot.caption.position = "plot",
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(size = 11),
    panel.grid.minor = element_blank()
  )
```
